### 2023-03-28 index.qmd
2nd

---
title: "Giovanna Lino"
about:
  template: jolla
  image: myphoto.jpeg
  links:
    - icon: linkedin
      text: Linkedin
      href: https://www.linkedin.com/in/giovannarlino/
    - icon: github
      text: Github
      href: https://github.com/giovannarlino
---

Hi, I'm Giovanna and I'm a student at the George Washington University. I like business and finance and enjoy keeping up with fashion and italian cuisine.

## Education

B.S. Degree in International Business and Finance from the George Washington University

## Skills

Reporting and Writing

Business Valuation

Marketing strategy