### 2023-04-10 portfolio.qmd

---
title: "Portfolio"
---

Hi there this is my portfolio page! Yay...

This is an example of one of my [works](https://www.linkedin.com) during my time at GW.