### 2023-04-10 papers.qmd

---
title: "Papers"
---

This is a paper I have written when I took the course called Introduction to International Politics (PSC-1003) at the George Washington University with Professor Downes in 2020. The argument revolves around the bargaining model of countries when they go to war and their negotiations methodologies and techniques. Furthermore, the paper explores the trade offs of those methodologies.

**The Bargaining Model**

War is defined as a "sustained, coordinated violence between political organizations" (Levy & Thompson, 2010, p. 5). As stated in Clausewitz (1996), "the political object is the goal, war is the means of reaching it, and means can never be considered in isolation from their purpose" (p. 87). However, war is not something states normally want. As cited in Fearon (1995), "war can be and often is a product of rational miscalculation" (p. 390), because rational states would normally avoid war due to its high costs. For each state, the value of war is "the share of the good it expects to win, minus the costs it expects to incur" (Frieden, Lake, & Schultz, 2019, p. 98). When states evaluate whether to go to war or to settle, they analyze the costs and values of entering a fight. Assuming that war is costly, states want to avoid it. Therefore, a bargaining range exists. This range is a set of agreements that makes each state at least as well off as it would be by going to war. However, since states live in an anarchic, self-help system there is no credible threat that can enforce states to settle, and without "such credible threat, war will sometimes appear the best option for states that have conflicting interests" (Fearon, 1995, p. 384). The central key for understanding the bargaining model of war is to understand why wars happen if they are so costly. To answer this question the bargaining model assumes five concepts: states are unitary actors that exist in anarchy; states pay the cost of war themselves; states do not prefer war for its own sake; states only care about getting the most valuable deal for themselves; states are rational and weigh costs and benefits. Considering those premises, there are three main reasons for war to occur and for a bargaining range to disappear.

Private information among the international system is one of the key factors that may lead states to war. Given the uncertainty that states face "bargaining strategies that entail a risk of war can be perfectly rational" (Frieden, Lake, & Schultz, 2019, p. 108). This is due to the lack of knowledge that states usually have of their opponent's military power and resolve to fight. When states have access to others' private information, the information needs to be analyzed very carefully, for states may have incentives to misrepresent them, leading their opponents' to overestimate their military power, and underestimate the other's, creating excessive optimism that may prove to be prejudicial in the future. Therefore, states can be misled by others that share misrepresented information in order to hide their weaknesses. States can indeed "exaggerate their true willingness or capability to fight, if by doing so they might deter future challenges or persuade the other side to make concessions. States can also have an incentive to conceal their capabilities or resolve, if they are concerned that revelation would make them militarily vulnerable or would reduce the chances for a successful first strike" (Fearon, 1995, pp. 395-396). Therefore, leaders can have various reasons to misrepresent private information in order to accomplish their goals, and that, in return, may cause the strike of a war. The article of Lake (2010/11) explained the rationalist explanation for the Iraq war using the bargaining theory. The conflict between the US and Iraq can be seen as due to the misrepresentation of information from Iraq. Saddam Hussein could not state the truth about his WMD capabilities to the US, even though he did not have developed any nuclear weapon, because his enemy, Iran, and other domestic opponents would see him as military weak and vulnerable. Moreover, Saddam believed that the US threat to invade Iraq was a bluff because the US lacked resolve, and would only launch a limited attack against the country, meaning that the US lacked credibility when communicating his threat to Saddam. On the other side, the Bush administration was committed to launching an attack against Iraq and did not see any possibility of settlement with the country, underestimating the costs of post-war governance in the US. The bargaining range between these countries disappeared since the US lacked credibility, Iraq had incentives to misrepresent information, and neither side wanted to know about different information that would have challenged each state's prior beliefs and slowed the path to war. In the end, Saddam hid the information regarding Iraq's nuclear weapons, resulting in his removal from power.

The second reason why a bargaining range may not exist relies on one of the most accepted theories that explain that the lack of a bargaining range between states is due to "the difficulty that states can have making credible promises not to use force to revise the settlement at a later date" (Frieden, Lake, & Schultz, 2019, p. 118). This is known as the commitment problem. There exists certain goods that can serve as a source of future bargaining power among states. Therefore, as stated in Frieden, Lake, & Schultz (2019), bargaining over such objects (will make the state) be reluctant to render itself more vulnerable to attack without credible promises that the other state will not exploit that vulnerability in the future (p. 118). This causes a preventive war to happen. As stated in Frieden, Lake, & Schultz (2019), this type of war occurs when there is a lack of balance of military capabilities between states because of factors external to the bargaining process, as a power shift in different rates of economic growth and development of new technologies (p. 120). This signifies that there is a rising state and a declining one. As a result, the weaker state may choose to fight now than to fight in the future when it will be worse off. For instance, consider a state A and a state B. When analyzing the possibility of war, "if B's expected decline in military power is too large relative to B's costs for war, then state's A inability to commit to restrain its foreign policy demands after it gains power makes preventive attack rational for state B" (Fearon, 1995, p. 406). A real life example that combines the concept of private information and commitment problem resulting in a preventive war is the War of Triple Alliance. As state through Weisiger (2013) article, Paraguay feared Argentina's and Brazil's approximation and conquest of Uruguay, stating that those states would threaten to blockade Paraguay's only outlet to the sea. Therefore, Paraguay launched a preventive war against Brazil. Lopez, Paraguay's president, could not commit to any settlement because of his fear of his inevitable decline. In addition, Brazilians believed that Lopez was lying about his reason for attacking the country, so they concluded that Lopez was untrustworthy, unable to commit to any settlement. As a result, Brazil saw the only way to peace as a change in Paraguay's regime by war. There was no bargaining range from either countries as Paraguay held a strong commitment problem. In addition, private misrepresented information also contributed to the downfall of Lopez. During the war, Paraguay only attacked Argentina due to the private information Lopez received that Argentinians would not follow war against the country in case he attacked Argentina. Therefore, Lopez overestimated his chances of winning and attacked Argentina as well. To his surprise, Lopez's private information was false, and Argentinians fought against Paraguay, contributing to its defeat.

Another concept that favors the idea that it is hard for states to credibly commit to a deal is the offensive advantage, through geography or military technology, that some states may have. These advantages help states strike a preemptive war, for when "first-strike advantages are large, both states must be given more from the peacetime bargain in order to allay the greater temptation of unilateral attack" (Fearon, 1995, p. 403). Therefore, states facing this situation are not concerned on making settlements with each other, but are instead focused on which state will strike first, and "negotiations in this context may be seen as nothing more than a ploy to delay the other side from mobilizing" (Frieden, Lake, & Schultz, 2019, p. 125). For instance, the Six Day War between Israel and the four Arab States (Egypt, Syria, Jordan, and Iraq) was a preemptive war. According to Frieden, Lake, & Schultz (2019), Egypt imposed a partial blockade on Israeli border with Syria in response to Israel's conflict with the border country (p. 125). Israel felt that Egypt would attack, and saw the opportunity of a first-strike advantage, even though Egypt did not plan to fight. This eliminated any possible bargaining range between states, for Israel now thought that the only way to solve the situation was to strike a first attack. Therefore, Israel launched a surprise attack on Egyptian air force, destroying Egypt's main offensive weapon, and beginning to control a large territory formerly held by Arabs.

The third and last reason why the bargaining range may disappear relates to the indivisibility of a good. Some goods may not be divided among states without losing their value. This is an all or nothing bargaining situation, and "if both states prefer war over getting nothing, then war becomes inevitable" (Frieden, Lake, & Schultz, 2019, p. 127). For instance, Jerusalem is a piece of land that "contains some of the holiest sites of Christianity, Islam, and Judaism, and historical, cultural, and religious significance unlike any other piece of territory in the world" (Frieden, Lake, & Schultz, 2019, p. 128). Therefore, Jerusalem stands as an obstacle to the resolution of conflicts in the Middle East. How to divide the territory and how to ensure that every religion has access to their holy sites is a question that is long for being resolved. That is one of the reasons why countries usually result to war when fighting for this land, for there is no way to 'divide' the territory for every religion without losing the territory's value. Although Jerusalem today is opened for every religion, each would prefer to have the territory for its own. Therefore, Jerusalem is a land that contributes to the religious/territorial conflict in the Middle East.

The bargaining model can also be applied to a current conflict in the world: the North Korea vs. US conflict. According to Frieden, Lake, & Schultz (2019), it is unknown what North Korea seeks to accomplish when developing nuclear weapons, and there exists different perspectives on how the US should deal with the country's approach to nuclear power (pp. 104-105). On the one hand, some people argue that is not possible to negotiate a bargain with North Korea due the state's commitment problems. The fact that nuclear weapons are a way to reunify Korean peninsula by force reinstates how valuable these weapons are to North Korea and how unlikely negotiations for disarmament will succeed. These people also argue that negotiations in this case serve only to buy North Korea more time to develop its nuclear program, relating to the concept of preemptive war. Another point they make is regarding North Korea's easiness to cheat to any commitment since the country has an autocratic secret regime, making it easy for it to misrepresent information about its nuclear disarmament. Therefore, North Korea is an untrustworthy country. Contributing to this distrust is North Korea's history of breaking deals with the US.

On the other hand, some people believe that North Korea develops nuclear weapons because of Kim Jong-Un's fear of being overthrown from power by the US. In that way, negotiations might work, but only if Washington commits not to seek regime change if North Korea lowers its defense, and commits to promote economic assistance and development in the North Korean regime and to reduce US military in Japan and South Korea. President Trump has been advised with different opinions on the matter. Some of his advisors believe that Trump and his allies should put pressure on North Korea to disarm, including having a limited military strike to convince Kim Jong-Un of the costs of defiance by a broader attack designed to eliminate his arsenal. Other advisors, however, believe that this aggressive measure is counterproductive and risky. If North Korea is interested solely in its own survival, then just the threat of retaliation by the US can deter North Korea from using nuclear weapons. In this conflict, the bargaining model is very controversial. Although, there are various set of possible deals between these two countries, one thing is certain, since both countries possess nuclear power, the possibility of war imposes too many costs for it to be gambled. For that reason, Trump is trying to find the most reasonable deal with North Korea to avoid any war from occurring. The reasons for this conflict sometimes to appear not to have any bargaining range is due to North Korea's commitment problems and misrepresentation of information. However, the consequences of not finding any settlement may result in a nuclear war. Therefore, there needs to exist a bargaining range, for negotiating with North Korea seems as the best diplomatic and rational approach for the resolution of the conflict.

The bargaining model does provide a credible explanation for the causes of war. However, it fails to explain certain aspects that may also lead or not to a conflict. First, the model does not consider postwar costs that may influence a state's decision of going to war or to settle. Second, the theory does not take into account the first image of analysis. Leaders may have motivated and unmotivated biases and specific personalities that may lead states not to have a bargaining range with others. Third, the model does not consider the domestic politics of states such as the regime type, special interests, military influence, and socio-economic problems that may make a country more risk accepting and make it go to war, or more risk-averse and make it prefer any settlement over war. At last, the bargaining model does not take into account multiple states in the brink of war, only two. Therefore, it does not explain how multiple states bargain, and the causes for the occurrence of war among them. As a result, although the bargaining model fails to address other relevant factors when analyzing the causes of war as postwar costs, cognitive biases of decision-makers, states' domestic politics, and conflicts among multiple states, the model provides a good set of explanations for the origins of conflict between two states.

\`\`\`