### 2023-04-10 portfolio.qmd

---
title: "Portfolio"
---

Hi there this is my portfolio page! Yay...

This [page](https://www.linkedin.com/in/giovannarlino/) showcases some of my works during my time at GW.