### 2023-04-17 virginiaelectionresults.qmd
another commit

---
title: "Virginia Election Project"
author: "Giovanna Lino"
execute:
  echo: true
format:
  html:
    self-contained: true
    code-tools: true
---

```{r setup, include=FALSE, warning=FALSE, message=FALSE}

library(tidyverse)
library(janitor)
library(httr)
library(jsonlite)
library(kableExtra)
library(here)
options(scipen = 999)
options(stringsAsFactors = FALSE)

# library(knitr)
# # print everything as paged ----
# knit_print.data.frame <- function (x, options, ...) {
#   rmarkdown::paged_table(x, options) |>
#     rmarkdown:::print.paged_df()
# }
# registerS3method("knit_print", "data.frame", knit_print.data.frame)

```

# Getting 2020 Presidential results

Data available here: <https://historical.elections.virginia.gov/elections/view/144567/>

A little column cleaning and we'll load in the data file.

```{r}

prez_2020 <- read_csv("processed_data/va_2020_prez_cleaned.csv")

```

Let's see what we have

```{r}

head(prez_2020)

```

Calculating percentage of the vote

```{r}

prez_2020 %>% 
  mutate(
    biden_pct = biden/total_votes_2021_prez
  )

```

Now let's do some rounding and move that decimal point

```{r}

prez_2020 %>% 
  mutate(
    biden_pct = janitor::round_half_up(biden / total_votes_2021_prez * 100, 1)
  )

```

Now trump too

```{r}

prez_2020 <- prez_2020 %>% 
  mutate(
    biden_pct = janitor::round_half_up(biden / total_votes_2021_prez * 100, 2),
    trump_pct = janitor::round_half_up(trump / total_votes_2021_prez * 100, 2)
  )

head(prez_2020)

```

# Getting 2021 Virginia governor results

Data now available from the state here: <https://historical.elections.virginia.gov/elections/view/147466>.

After the election and for few years after, was only available in JSON format.

<https://results.elections.virginia.gov/vaelections/2021%20November%20General/Json/Governor.json>

But site location doesn't exist anymore. We can see it using the wayback machine here:

<https://web.archive.org/web/20220806120753/https://results.elections.virginia.gov/vaelections/2021%20November%20General/Json/Governor.json>

We'll use an archived file of the in JSON to learn about how to deal with it that format.

```{r}

jsonfile <- "raw_data/va_gov_json_archived.json"

#using jsonlite package function fromJSON()
this.content <- fromJSON(jsonfile)

#dataframe from just the 6 content 
content_df <- as.data.frame(this.content[[6]])

```

Where are candidates themselves? They are "nested" inside. We'll use `unnest()`to expand things.

```{r}
#unnest
results <- content_df %>%
  unnest(cols = Candidates)

head(results)

```

Unnest again on locality

```{r}

results <- results %>%
  unnest(cols = Locality)

head(results)
```

Great.

We'll give it a new name to make it easier for us and clean the column names, remove extraneous ones.

```{r}

gov_2021 <- results %>% 
  clean_names() %>% 
  select(-precincts_reporting,
         -precincts_participating,
         -last_modified,
         -ballot_order)

head(gov_2021)

```

What's the issue we may have here?

This is actually tidy data which is good - but for comparing the candidates we'd want them on the same row.

## Reshaping

Enter pivot_wider().

We'll get rid of everything we don't need first.

```{r}

gov_2021 <- gov_2021 %>% 
  filter(ballot_name %in% c("Glenn A. Youngkin", "Terry R. McAuliffe")) %>% 
  select(-locality_code,
         -political_party)

gov_2021

```

Now we'll do the spreading out to reshape. One value for each locality

```{r}

gov_2021_wide <- gov_2021 %>% 
  pivot_wider(names_from = ballot_name, values_from = c(votes, percentage))

gov_2021_wide

```

Nice.

This is giving us some pretty long column names. we can change them after the fact using `rename()`. But first let's clean the names to make it easier.

```{r}

gov_2021_wide <- gov_2021_wide %>% 
  clean_names()

head(gov_2021_wide)
```

Now let's rename, and we'll use similar names to what we had earlier in our 2021 results.

```{r}

gov_2021_wide <- gov_2021_wide %>% 
  rename(
    youngkin = votes_glenn_a_youngkin,
    mcauliffe = votes_terry_r_mc_auliffe,
    pct_youngkin = percentage_glenn_a_youngkin,
    pct_mcauliffe = percentage_terry_r_mc_auliffe
  )

head(gov_2021_wide)

```

Bingo.

There's still one potential issue here. Can you see it?

The percentage columns are actually text values, not numbers. And they have that `%` sign in the text too. Let's fix that using a handy function from the readr package, `parse_number()`.

```{r}

gov_2021_wide <- gov_2021_wide %>% 
  mutate(
    pct_youngkin = readr::parse_number(pct_youngkin),
    pct_mcauliffe = readr::parse_number(pct_mcauliffe)
  )

head(gov_2021_wide)
```

Perfect. Problem solved.

# Putting things together - joining

Now we want to actually join things up.

Can we do that? Let's check out our two tables

```{r}

gov_2021_wide

```

```{r}

prez_2020

```

Few quick things to deal with... first let's get rid of the total votes in 2021, we don't need that anymore. But also the locality isn't the same case as in our governor's results, so let's uppercase it.

```{r}

prez_2020 <- prez_2020 %>% 
  mutate(
    locality = str_to_upper(locality)
  ) %>% 
  select(-total_votes_2021_prez)

prez_2020

```

Great. Now we're all ready to join. Or are we?

Look closely at the record counts for our two tables. It looks like the presidential results has an extra record perhaps? Let's investigate.

We can use `anti_join()` to see what's in one table but not the other.

```{r}

anti_join(prez_2020, gov_2021_wide, by = c("locality" = "locality_name"))

```

There's a total row in there! Well we don't want that. Let's take it out.

Also though notice a second county didn't match...want to be that might be because it's differently worded? Let's see by doing the anti-join in reverse now.

```{r}

anti_join(gov_2021_wide, prez_2020, by = c("locality_name" = "locality"))

```

Yep that's it. So let's fix both of those things in the prez table.

```{r}

prez_2020 <- prez_2020 %>% 
  filter(locality != "TOTALS") %>% 
  mutate(
    locality = str_replace(locality, "KING AND QUEEN", "KING & QUEEN")
  )

```

Now we're ready. Let's join.

```{r}

joined_vacomparison <- inner_join(prez_2020, gov_2021_wide, by = c("locality" = "locality_name"))

head(joined_vacomparison)

#save results to file for next step
saveRDS(joined_vacomparison, here("processed_data", "joined_vacomparison.rds"))
write_csv(joined_vacomparison, here("processed_data", "joined_vacomparison.csv"))

```

Alright! It worked.

## Comparing gov vs. prez results

Now that things are join, let's actually go ahead and start making columns to compare the two elections and how the candidates did this time compared with last time.

Where should we go from here....? Give it a shot...

```{r}

joined_vacomparison %>% 
  mutate(
    mc_overperform = pct_mcauliffe - biden_pct,
    mc_overperform_worsethan5 = if_else(mc_overperform < -5, "YES", "NO") #created a flag column

  )

```

```{r}

#save results to file for next step
saveRDS(joined_vacomparison, here("processed_data", "joined_vacomparison.rds"))
write_csv(joined_vacomparison, here("processed_data", "joined_vacomparison.csv"))

```

# **My Analysis**

```{r}

#load saved joined data file from previous script
joined_vacomparison <- readRDS(here("processed_data", "joined_vacomparison.rds"))

```

# Comparing Virginia Gov vs. President Victory

```{r}

head(joined_vacomparison)

```

```{r}
#what is the average democratic percentage for both presidential and government election in Accomack County?

joined_vacomparison %>% 
  filter(locality == "ACCOMACK COUNTY" ) %>% 
  group_by(biden_pct, pct_mcauliffe) %>% 
  summarise(average = mean(biden_pct)) %>% 
  arrange(desc(average))

```

# **Possible questions to be asked based on the data:**

1.  Create a map of the republican vote in all of the counties for both president and governor.

2.  Analyze the counties with the biggest democratic victory.

3.  Analyze the counties with the biggest republican victory.

4.  Was there a county in which Biden won for which the democratic governor didn't?