### 2023-05-03 about.qmd
annother commit

---
title: "About Me"
---

Hello! My name is Giovanna Romariz Lino. I am 22 years old and I am originally from Londrina, Brazil. I have just graduated from George Washington University with a Bachelor of Science degree in Business with a double concentration in International Business and Finance and a minor in Journalism and Mass Communications. During university, I have held leadership roles in the Brazilian Student Association, Alpha Kappa Psi Business Fraternity, the GW Socially Responsible Fund, among others. I have spent 4 years living in Washington D.C., and I have also studied abroad in Milan, Italy, at Bocconi University during Spring, 2022. During my life, I have also lived in Canada, and done summer exchanges in the United States while I was in high school. I speak fluent English, Portuguese, intermediate Spanish, and basic Italian. I am an only child and my parents still live in Brazil.

During my time at university, I have interned in more than four companies in Brazil focused on finance and marketing. I have also participated in consulting projects in Washington D.C. My main focuses of work following graduation are on consulting and finance. Now, I am moving to New York in July, 2023, and joining Alphasights as an Associate working under their Consulting department.

If you want to know more about my professional experience, you can read my [resume.](https://docs.google.com/document/d/1cv7g5l-lgJfpHSnO49QV12rrD57OgwzQ/edit?usp=sharing&ouid=116938803250525408172&rtpof=true&sd=true)