### 2023-05-03 index.qmd

---
title: "Giovanna Lino"
about:
  template: jolla
  image: myphoto.jpeg
  links:
    - icon: linkedin
      text: Linkedin
      href: https://www.linkedin.com/in/giovannarlino/
    - icon: github
      text: Github
      href: https://github.com/giovannarlino
---

Hi, welcome to my webiste! I'm Giovanna and I'm a student at the George Washington University.

## Education

B.S. Degree in Business with a double concentration in International Business and Finance and a minor in Journalism and Mass Communications from the George Washington University in Washington D.C.

## Skills

Financial analysis and valuation

Reporting and writing

Inbound and Outbound Marketing Strategy

## Interests

Business podscasts, italian cuisine and fashion.