### 2023-05-03 newsstory.qmd

---
title: "News Story"
author: "Giovanna Lino"
---

# Voter Suppression in Texas

*September, 2022*

The expression 'Fraudulent Election' has risen into prominence once again as former president Donald Trump has questioned the veracity of the results of the 2020 Presidential Election, when he lost to current Democratic President Joe Biden. Texas, a traditional Republican state governed by Greg Abbott, answering to Trump's claims, has passed a Senate Bill tightening state election laws in 2022, which was retaliated with claims of voter suppression in the state of Texas.

In light of Abbott's third special season in government, Trump has pressured the current Texas Governor to pursue an election audit in most of the state's counties, claiming voter fraud and irregularities in the 2020 elections. Eight and a half hours later, Abbott's administration pursued a forensic auditing process on four of the state's major counties - Collin, Dallas, Harris (Houston included), and Tarrant, showcasing the influence that Trump still has on the state's Republican party. However, the vote reassessment did not uncover any impropriety, and there is no evidence of fraud in the state during the election period.

In response to Abbott's actions, Trump wrote a letter claiming that Texans \"do not trust (Abbott's) election system, and they want (his) leadership on this issue, which is the number one thing they care about.\"

In September 2021, Abbott signed a senate bill (SB1), which bans 24 hour voting, drive-thru voting, and imposes difficulties on voting for mail-in-ballot, people with disabilities and marginalized groups. Those measures also protect partisan poll watchers and set limits to those who help other voters, with their commitment of swearing under oath.

"What makes this bill and similar ones Republicans are pushing across the country even more un-American is that Republicans are using the 'Big Lie' about the 2020 election as a pretext to support them. The reality is that these bills have nothing to do with election integrity or security, but rather are discriminatory measures making it harder for all people to vote. These bills will have a disproportionate impact on communities of color," stated Eric Holder, the US attorney general under former president Barack Obama.

Democratic candidate for Governor, Beto O'Rourke, has argued in Newsweek that "voter suppression is on purpose and by design, working exactly as Greg Abbott intended". He criticized Abbott over reports that election officials have sent back unapproved mail-in-ballot to voters due to identification problems. Mail-in-ballots affects voters over 65 years of age who cannot provide a correct driver's license or social security number, military veterans, and overseas voters.

Furthermore, The New York Times found that after the primaries held in the beginning of 2022, Harris county, the most populous county in Texas, held that "areas with large Black populations were 44 percent more likely to have ballots rejected than heavily white areas." Jen Ramos, the political expert for JOLT Action, a progressive club that works with Texan Latinos, told Newsweek that ballot rejections in El Paso, a city with 85% Latinos, was 15 percent during primaries in 2022. The implementation of SB1 has resulted in civil rights groups and democrats filing lawsuits against the new legislation claiming it as an attempt of voter suppression for marginalized communities.

\"Instead of working on issues that actually matter, like protecting school kids from Covid or fixing our failing electrical grid, Abbott is focused on rigging our elections and implementing extreme, right-wing policies,\" said O\'Rourke to The New York Times.

The Gubernatorial Election of Texas in 2022 will be held with no online voter registration on Election Day; stricter ID policies especially for Latinos, communities of color, and people with disabilities, who are less likely to have an acceptable form of identification on hand; no 24 hour election period accommodation; increased limitations for voter's assistants; and stricter criminal penalties.

Abbott's administration continues to claim that the bill facilitates instead of suppresses the voting process, but, with all these obstacles, it is worth questioning the real agenda behind the SB1.

\

\